Big data' og politologisk datavidenskab

### 2020-03-31 draft_v01.Rmd
some comments & added grafs to ethics section

@@

@@

Princippet om at big data ikke er lavet for samfundsforskningens skyld gælder også for API'er. Det egentlige formål for de fleste API'er er at dele data på tværs af kommercielle platforme. For eksempel er det API'er der muliggør at et online-medie kan vise hvilke af ens egne Facebook-venner der har 'liket' en specifik artikel, fordi avisen kan tilgå data om læserens Facebook-netværk gennem Facebooks API. Men mange sociale netværk stiller meget righoldige data til rådighed for forskere gennem API'er. For eksempel bruger @hjorth2019, som studerer fordelingen af online misinformation, Twitters API til at indsamle data om ca. 13 millioner følgere af ca. 10.000 Twitter-konti. Offentlige myndigheder stiller også i stigende grad data til rådighed gennem API'er. Eksempelvis stiller Folketinget data om medlemmer, forhandlinger og lovarbejde til rådighed gennem en API.

@@

Big data has applications in political science in empirical studies of all types, from description to explanation, prediction, and causal studies. Although these applications are only beginning across the social sciences, the past 15 years have provided enough experience that we can already point to one strong finding that can always guide us when we apply big data in our work: research design still matters [see @toshkov16, p. 13; @CG15]. Here we discuss aspects of research design that deserve special attention when working with big data, namely, biases in the data and the definition of the unit of analysis.

@@

Perhaps the most heady idea about big data was expressed best in a, now infamous, article in *Wired* magazine by Chris @anderson08 -- the idea that big data eliminates sampling problems because *n*=all. That is, one can analyze *all of the data*. The article has met with a good deal of attention and pushback in the years since it was published. In early 2020, it boasts more than 2.000 citations on Google Scholar. To put that in perspective, Maurice Duverger's famous book *Political Parties*, the origin of the eponymous Duverger's Law, was published in 1959 and counts just over 7.000 citations.

As a concept, *n*=all implies that big data can be taken at face value. Its patterns reveal a complete picture of human behavior. Unfortunately, the selection problem is still with us. It arises whenever observations enter your data for reasons that systematically relate to the outcome variable. @AP08 illustrate the problem with an example from the United States' National Health Interview Survey, in which individuals are asked to rate their health on a scale. Perhaps unsurprisingly, they find that individuals in hospitals consistently rate their health as worse than those out of hospital. Then they pose an interesting question: should we conclude that hospitals make people sicker? The answer is, of course, no. We know that because we already understand that individuals in hospital *selected* into being there precisely because they were sick. Failing to recognize that would result in sampling bias.

@@

Big data confronts us with new sources of bias and confounding. Perhaps the most famous example for social scientists is that of Google Flu Trends (GFT). GFT was a Google project launched in 2008 that predicted regional flu epidemics from fine-grained data on users' Google queries about likely flu symptoms [@Getal2009]. After initially receiving attention for its impressive accuracy, GFT's predictive performance began to diminish over time until--by the time the project was shuttered in 2015--for years it had produced inaccurately high forecasts5 sometimes as great as double the number of flu reported by the U.S. Centers for Disease Control and Prevention [@harford14]. The reasons for GFT's collapse are instructive for understanding both why research design still matters and what new headaches come with big data.

Several things contributed to GFT's problems, but the most noteworthy seems to have arisen after Google adjusted its main search service [see @LKKV14]. Like any organization that produces or processes big data, Google changes its service or its algorithms from time to time. In 2011 and 2012, Google added functions to its search service that suggested additional search terms to users, based on their original search terms. It even suggested possible diagnoses for search terms that involved disease symptoms. The upshot was an apparent feedback loop: users of Google search were nudged to refine their searches for symptoms and to seek out information on the flu, intensifying the signals that GFT relied on to predict flu outbreaks. There is more to GFT's story, of course, but this part of it points out some special features of big data researchers must consider to avoid sampling bias. As @salganik17 puts it, big data is *algorithmically confounded*, *dirty*, and *drifting*.

@@

By "dirty," @salganik17 means that big data also contains false positives in the form of computer behavior that is mistaken for human behavior. For example, at least a portion of Google's additional search traffic around flu syptoms may have been purely accidental, caused only by the search engine's suggestions. Smart phone users will recognize this problem as an autocorrect fail. An example that occassionaly makes headlines would be the presence of large numbers of "bots" on social media platforms. Bots are accounts that are not operated by human users, but rather are simply software programs executing automated behaviors in the form of liking, following, or commenting. Despite Twitter and other platforms' occasional purges of bot accounts, an unspecified and probably high number of bots persist. Any analysis of social media data must think carefully about how to clean the data to eliminate as much bot behavior as possible.

@@

Finally, develop a habit of running sanity checks, even when you do not anticipate problems and cannot formulate an explicit assumption. Since big data is, well, big, it is not realistic to look at the numbers or text in your data and expect to see anything useful. We do this by drawing pictures. Make histograms, scatterplots, and other plots of your variables and be sure the patterns make sense. When you discover a strange pattern, investigate it until you understand it. You may very well uncover hidden drift or the tell-tale patterns of dirty data caused by bots or other automated interference.

@@

In the next section, we introduce several popular and useful methods for working with big data. As we will see, these methods specifically associated with big data have in common that they are actually measurement tools. We will look at methods for labeling data, discovering patterns, and condensing large amounts of data into simpler, smaller data. It is typical for political scientists applying big data that their "big" data set of a million tweets or thousands of laws condenses into a data set of, for example, thousands of politician-month observations or hundreds of government-year observations.

The workhorse method of observational, quantitative hypothesis-testing in political science, even when working with big data, remains to build and estimate appropriate statistical models with appropriate control variables. Thus, we will focus on the methodological choices that distinguish analyses using big data from the more familiar social science workflow of building a purpose-built data set arranged at the correct level of analysis. Readers who are familiar with qualitative methods will be more familiar with this pattern. For example, lengthy interviews or wide-ranging case studies are often condensed into a summary coding of theoretical concepts.

This raises a point worth emphasizing about big data methods in political science. Although methods for working with big data are, necessarily, computational they do not necessarily mean that researchers must apply quantitative methodology in their overall research design.

<!-- Consider an example. Suppose you  data on the Twitter posts of Danish politicians.  -->

## measurement stuff

<!-- here we transition to the discussion of units of analysis and jusifying the focus on measurement -->

- role of big data in the research design
    - uncommon to directly analyze
    - more often: measurement

@@

>>>>

<!-- ML: Should we say "en række specifikke tekniske værktøjer" Seems more like we'll present 'classes' of specific methods or 'families' of specific methods with references for further reading -->

@@

<!-- ML: This seems like a good place to mention a thing that really frustrates me! The robots.txt for Denmark's Supreme Court's decision database disallows all scraping: http://domstol.fe1.tangora.com/robots.txt I haven't tried poking the bear and doing it anyway, but I often use this as an example of when researchers have to make explicit trade-offs in their data collection. If the Supreme Court were to refuse a request to share its data, I think it would be thoroughly **ethical** to violate these scraping guidelines. But, at what cost to the researcher? -->

@@

Princippet om at big data ikke er lavet for samfundsforskningens skyld gælder også for API'er. Det egentlige formål for de fleste API'er er at dele data på tværs af kommercielle platforme. For eksempel er der API'er der muliggør at et online-medie kan vise hvilke af ens egne Facebook-venner der har 'liket' en specifik artikel, fordi avisen kan tilgå data om læserens Facebook-netværk gennem Facebooks API. Men mange sociale netværk stiller meget righoldige data til rådighed for forskere gennem API'er. For eksempel bruger @hjorth2019, som studerer fordelingen af online misinformation, Twitters API til at indsamle data om ca. 13 millioner følgere af ca. 10.000 Twitter-konti. Offentlige myndigheder stiller også i stigende grad data til rådighed gennem API'er. Eksempelvis stiller Folketinget data om medlemmer, forhandlinger og lovarbejde til rådighed gennem en API.

@@

Big data has applications in political science in empirical studies of all types, from description to explanation, prediction, and causal studies. Although these applications are only beginning across the social sciences, the past 15 years have provided enough experience that we can already point to one strong finding that can always guide us when we apply big data in our work: research design still matters [see @toshkov16, p. 13; @CG15]. Here we discuss aspects of research design that deserve special attention when working with big data, namely, several biases and the definition of the unit of analysis.

@@

Perhaps the headiest idea about big data was expressed best in a, now infamous, article in *Wired* magazine by Chris @anderson08 -- the idea that big data eliminates sampling problems because *n*=all. That is, one can analyze *all of the data*. The article has met with a good deal of attention and pushback in the years since it was published. In early 2020, it boasts more than 2.000 citations on Google Scholar. To put that in perspective, Maurice Duverger's famous book *Political Parties*, the origin of the eponymous Duverger's Law, was published in 1959 and counts just over 7.000 citations.

<!-- ML: The citation counts here are not explicitly necessary, I just enjoy pointing out how many people have dunked on Chris Anderson's article. -->

As a concept, *n*=all implies that big data can be taken at face value. Its patterns reveal the complete picture of human behavior. Unfortunately, the selection problem is still with us. It arises whenever observations enter your data for reasons that systematically relate to the outcome variable. @AP08 illustrate the problem with an example from the United States' National Health Interview Survey, in which individuals are asked to rate their health on a scale. Perhaps unsurprisingly, they find that individuals in hospital consistently rate their health as worse than those out of hospital. Then they pose an interesting question: should we conclude that hospitals make people sicker? The answer is, of course, no. We know that because we already understand that individuals in hospital *selected* into being there precisely because they were sick. Failing to recognize that would result in sampling bias.

@@

Big data confronts us with new sources of bias and confounding. Perhaps the most famous example for social scientists is that of Google Flu Trends (GFT). GFT was a Google project launched in 2008 that predicted regional flu epidemics from fine-grained data on users' Google queries about likely flu symptoms [@Getal2009]. After initially receiving attention for its impressive accuracy, GFT's predictive performance began to diminish over time until--by the time the project was shuttered in 2015--for years it had produced inaccurately high forecasts sometimes as great as double the number of flu cases reported by the U.S. Centers for Disease Control and Prevention [@harford14]. The reasons for GFT's collapse are instructive for understanding both why research design still matters and what new headaches come with big data.

Several things contributed to GFT's problems, but the most noteworthy seems to have arisen after Google adjusted its main search service [see @LKKV14]. Like any organization that produces or processes big data, Google changes its services or its algorithms from time to time. In 2011 and 2012, Google added functions to its search service that suggested additional search terms to users, based on their original search terms. It even suggested possible diagnoses for search terms that involved disease symptoms. The upshot was an apparent feedback loop: users of Google search were nudged to refine their searches for symptoms and to seek out information on the flu, intensifying the signals that GFT relied on to predict flu outbreaks. There is more to GFT's story, of course, but this part of it points out some special features of big data researchers must consider to avoid sampling bias. As @salganik17 puts it, big data is *algorithmically confounded*, *dirty*, and *drifting*.

@@

By "dirty," @salganik17 means that big data also contains false positives in the form of computer behavior that is mistaken for human behavior. For example, at least a portion of Google's additional search traffic around flu syptoms may have been purely accidental, caused only by the search engine's suggestions. Smart phone users will recognize this problem as an autocorrect fail. An example that occassionaly makes headlines would be the presence of large numbers of "bots" on social media platforms. Bots are accounts that are not operated by human users, but rather are simply software programs executing automated behaviors in the form of liking, following, or commenting. Despite Twitter and other platforms' occasional purges of bot accounts, an unspecified and probably high number of bots persist. Any analysis of social media data must think carefully about how to clean the data to avoid analyzing bot behavior as if it were human behavior.

@@

Finally, develop a habit of running sanity checks, even when you do not anticipate problems and cannot formulate an explicit assumption. Since big data is, well, big, it is not realistic to look at the numbers or text in your data and expect to see anything useful. We do this by drawing pictures. Make scatterplots, histograms, and other plots of your variables and be sure the patterns make sense. When you discover a strange pattern, investigate it until you understand it. You may very well uncover hidden drift or the tell-tale patterns of dirty data caused by bots or other automated interference.

@@

In the next section, we introduce several popular and useful methods for working with big data. As we will see, these methods specifically associated with big data have in common that they are actually measurement tools. We will look at methods for labeling data, discovering patterns, and condensing large amounts of data into simpler, smaller data. It is typical for political scientists applying big data that their "big" data set of millions of tweets or thousands of laws condenses into a data set of, for example, thousands of politician-month observations or hundreds of government-year observations.

Even when working with big data sets, the workhorse method of observational, quantitative hypothesis-testing in political science remains to build and estimate appropriate statistical models with appropriate control variables. Thus, we will focus on the methodological choices that distinguish analyses using big data from the more familiar social science workflow of building a purpose-built data set arranged at the correct level of analysis. Readers who are familiar with qualitative methods will be more familiar with this pattern. For example, lengthy interviews or wide-ranging case studies are often condensed into a summary coding of theoretical concepts.

This raises a point worth emphasizing about big data methods in political science. Although methods for working with big data have to be computational, they do not necessarily mean that researchers must apply quantitative methodology in their overall research design. Descriptive and qualitative research designs can equally well apply the methods we discuss below.

@@

A final ethical concern to consider when applying big data relates to transparency and human bias. As we saw above, methods for making measurements from big data take their results directly from the data. Supervised classification learns associations between features and the target outcome, while unsupervised methods mine multidimensional associations in the data to return scales, clusters, etc. It is valuable to keep in mind that simply learning the associations in data is not neutral because data are not neutral. They are produced in the real world, complete with all of its biases: from the quirks of human perception to systematic racial prejudice or social inequality.

Uncritically learning associations in data can mean doubling down on negative aspects of the real world. For example, @oneill16 tells the story of how credit card companies in the U.S. market their products online to consumers, based on something called an *e-score*. The e-score is calculated from big data, including a potential customer's browsing history, location, and other data points. As a result, for example, people browsing the Internet from poorer areas of the country, with higher loan default rates, would generally receive a worse e-score. Lower e-scores mean the person sees advertisements for credit cards with less favorable terms, reflecting their apparent risk for the card issuer. This pattern contributes to deepening economic inequality and even racial disparities in wealth when neighborhoods are strongly racially segregated.

Any algorithm that learns from patterns in real-world data on human behavior can and will pick up associations of this nature -- legacies of inequality, racial and gender prejudice, etc. This is particularly relevant for the study of political science and public administration. The state makes life-altering decisions each day, for example, social workers help decide whether to remove children from potentially abusive homes into protective state custody or and case workers help match unemployed job-seekers to employment opportunities. If one were to apply a classification model, trained on historical data, to assist these processes, that model would inherit all the systematic biases or inequalities that may have characterized that system.

This is where transparency is crucial. Whether using big data for research or applying it in business or government, when issues like fairness or prejudice matter, it is crucial that we examine our models for these types of biases. This means slowing down, examining a model's results for bias, and being transparent about what we find before rushing to apply it. Many methods applied to big data are difficult or impossible to directly examine unlike, for instance, the coefficients of a regression model [@blackbox]. Therefore, it is up to the analyst to probe the model for evidence of bias by studying what the model predicts or what measurements it returns under a variety of circumstances. Algorithmic transparency is an active research field in computer science, but unfortunately right now there is no methodological trick that can substitute for carefully and systematically examining our models' performance.
